Refinements for open automata

### 2023-04-13 2021-Refinement/paper/refinementarticleV2.tex
Examples

* example

@@

* global rewriting

@@

\TODO{Debut Exemple-1}
\newpage
\begin{example} As a running example, we consider a classical producer-consumer pair  interacting through FIFO buffer,  referred to as \texttt{prod-cons}.  Fig. \ref{Fig:Architect} reflects the overall structure of the system involving a producer process, a consumer process and an orchestrator that coordinates  their activities.

\begin{figure}[h]
 \centering
   \includegraphics[width=.70\textwidth]{Figures/Architecture.pdf}
   \caption{Structure of the example. Each box corresponds to a process whose ports are the actions it can perform. The  actions observable by the environment are $\texttt{push}$, which indicates the enqueuing of an element, $\texttt{pop}$ which indicates the dequeuing, and $\texttt{print}$ which indicates the outing of results. \label{Fig:Architect}} 
\end{figure}

The open automaton  modelling the behaviour of such a system  using an  unbounded circular/ring buffer is depicted in Fig. \ref{Fig:RefineOA}.  The automaton has a single state with  two holes: $\texttt{P}$ and  $\texttt{Q}$ reflecting the processes to manage the producer and consumer respectively.  The two pointers $\texttt{l}$ and $\texttt{f}$  indicate resp. the next available position for  enqueuing an element  and the position that contains the next element to be dequeued %\footnote{The  arrow  without any state of origin is a graphical convention to indicate that the pointers are initialized at the beginning}. 
Whenever the buffer agrees with $\texttt{P}$ on the acquisition of an element, it enqueues it. Similarly, whenever it agrees with $\texttt{Q}$ on the retrieval, it dequeues an element. However, whenever $\texttt{Q}$ produces an item, it is exposed as an external observable  $\texttt{print}$ action. Whereas, when both processes do their internal computation, it is exposed externally as unobservable action $\tau$. 
\begin{figure}[h]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/FIFORing.pdf}
   \caption{Open Automaton for the \texttt{prod-cons} system using FIFO circular buffer.
    \label{Fig:RefineOA}} 
\end{figure}

\end{example}

\medskip
\TODO{Fin Example-1}

@@

\TODO{Debut -- Exemple 2}

\begin{example} Examples of processes that can fill the holes of the considered automaton can also modelled by open automata as shown is  Fig. \ref{Fig:procandcons}.
 \begin{figure}[h]
 \centering
   \includegraphics[width=.35\textwidth]{Figures/P-proc.pdf}\hfill 
   \includegraphics[width=.45\textwidth]{Figures/Q-proc.pdf}
   \caption{(Left) An example of a \texttt{producer}.  It produces one item at a time and puts it to the buffer.  (Right)  An example of a \texttt{consumer}. It gets an item from the buffer, does its work and returns the result. \label{Fig:procandcons}}
\end{figure}

The open automaton on Fig. \ref{Fig:ComposeOA} is the composition of the specification on Fig. \ref{Fig:RefineOA} and the producer on Fig. \ref{Fig:procandcons}. The composition produces two states (le product of the states of both automata), the transitions from one state to another come from  the synchronisation  of the transitions of the encompassing automaton including the hole \texttt{P} and those of the producer filling the hole. Transitions involving \texttt{Q} remain unchanged.

\begin{figure}[h]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/compos-FIFO-Pro.pdf}
   \caption{Open Automaton of the \texttt{prod-cons}[\texttt{P}/\texttt{producer}]. 
   \label{Fig:ComposeOA}} 
\end{figure}
\end{example}

\TODO{Fin -- Exemple 2}

@@

\TODO{Debut -- Example 3}
\begin{example} 
To illustrate the refinement simulation of open automata, consider a variation on the \texttt{prod-cons}.  We suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. The  corresponding behaviour  is given by  an open automaton  depicted in Fig. \ref{Fig:SpecOA}. 
\begin{figure}[h]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{Open Automation of the \texttt{prod-cons} system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that the automata on Fig.\ref{Fig:RefineOA} is a refinement of this automaton. $R=\{(r_0,s_0) \mapsto True, (r_0,s_1) \mapsto True \}$ is a simulation.
\end{example}
\TODO{Fin}

@@

>>>>

* example a terminer

@@

* global rewriting (partly done)

@@

\begin{example}[\texttt{prod-cons}]\label{ex:fifo-system} As a running example, we consider a classical producer-consumer pair  interacting through FIFO buffer,  referred to as \texttt{prod-cons}.  Fig. \ref{Fig:Architect} reflects the overall structure of the system involving a producer process, a consumer process and an orchestrator that coordinates  their activities.

\begin{figure}[!ht]
 \centering
   \includegraphics[width=.70\textwidth]{Figures/Architecture.pdf}
   \caption{Structure of the example. Each box corresponds to a process whose ports are the actions it can perform. The  actions observable by the environment are $\texttt{push}$, which indicates the enqueuing of an element, $\texttt{pop}$ which indicates the dequeuing, and $\texttt{print}$ which indicates the production of results. \label{Fig:Architect}} 
\end{figure}

The example uses several kinds of data. Variable $m$ holds a message, message types do not need a concrete definition here. We additionally use arrays of messages with a syntax of the form $M[l]$ for array accesses; $M$ is an array of $N$ elements, from $0$ to $N-1$. Finally we use addition and modulo operation ($\%$) on integers.

The open automaton  modelling the behaviour of such a system  using an  unbounded circular/ring buffer is depicted in Fig. \ref{Fig:RefineOA}.  The automaton has a single state with  two holes: $\texttt{P}$ and  $\texttt{Q}$ that are the two interacting processes.  $\texttt{l}$   indicates the next available position for  enqueuing an element  and  $\texttt{f}$ is the position that contains the next element to be dequeued. %\footnote{The  arrow  without any state of origin is a graphical convention to indicate that the pointers are initialized at the beginning}. 
The buffer reacts to a push from $\texttt{P}$ and enqueues it. Similarly, whenever it  $\texttt{Q}$ pops an element, it dequeues it. Additionally, whenever $\texttt{Q}$ produces an item, it is exposed as an external observable  $\texttt{print}$ action. 
When any process do its internal computation, it is exposed externally as unobservable action $\tau$. 
\begin{figure}[!tb]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/FIFORing.pdf}
   \caption{Open Automaton for the \texttt{prod-cons} system using FIFO circular buffer.
    \label{Fig:RefineOA}} 
\end{figure}
\qed
\end{example}

@@

\begin{example} 
 Fig. \ref{Fig:procandcons} shows a producer automaton  and a consumer  automaton that can be used  to fill the holes $P$ and $Q$ of \texttt{prod-cons} defined in Example~\ref{ex:fifo-system}.
 \begin{figure}[!ht]
 \centering
   \includegraphics[width=.35\textwidth]{Figures/P-proc.pdf}\hfill 
   \includegraphics[width=.45\textwidth]{Figures/Q-proc.pdf}
   \caption{(Left) A \texttt{producer}.  It produces one item at a time and pushes it.  (Right)  A \texttt{consumer}. It pops an item, does some work and pushes the result. \label{Fig:procandcons}}
\end{figure}

The open automaton on Fig. \ref{Fig:ComposeOA} is the composition of the system in Fig.~\ref{Fig:RefineOA} and the producer in Fig. \ref{Fig:procandcons}. The composition consists of two states (the product of the states of both automata). The transitions from one state to another come from  the synchronisation  of the transitions of the encompassing automaton with those of the producer filling the hole $P$. Transitions involving \texttt{Q} remain unchanged. \qed

\begin{figure}[!ht]
 \centering
   \includegraphics[width=.99\textwidth]{Figures/compos-FIFO-Pro.pdf}
   \caption{Open Automaton of the \texttt{prod-cons}[\texttt{P}/\texttt{producer}]. 
   \label{Fig:ComposeOA}} 
\end{figure}
\end{example}

@@

\begin{example} 
To illustrate the refinement simulation of open automata, consider a variation on the \texttt{prod-cons}.  We suppose that  the two processes  \texttt{P} and  \texttt{Q} communicate through a one-place  buffer. The  corresponding behaviour  is given by  an open automaton  depicted in Fig. \ref{Fig:SpecOA}. 
\begin{figure}[!t]
 \centering
   \includegraphics[width=.90\textwidth]{Figures/FIFOpen2.pdf}
   \caption{Open Automation of the \texttt{prod-cons} system using one-place  buffer.  \label{Fig:SpecOA}}
\end{figure}
We can easily check that this automaton simulates the one on Fig.\ref{Fig:RefineOA}.
The simulation relation is $R=\{(r_0,s_0) \mapsto l=f, (r_0,s_1) \mapsto f=l+1\% N \}$ is a simulation. \qed
\end{example}

@@

\texttt{prod-cons[P/producer] < simple-prod-cons[P/producer]}
